---
title: "Biocrusts in Kalahari Sands include spatially structured fungal communities which correlate with bacterial communities"
output:
  html_document:
    toc: yes
---

```{r setup, echo=FALSE, cache=FALSE}
options(width = 400)
# create folders for output of tables and figures
if (!dir.exists("tables")) {dir.create("tables")}
if (!dir.exists("figures")) {dir.create("figures")}
if (!dir.exists("figures_extra")) {dir.create("figures_extra")}

tidy_phyloseq <- function(my_phyloseq){
  # Fix ranks. 
  colnames(tax_table(my_phyloseq)) = c("Kingdom", "Phylum", "Class", "Order", "Family", "Genus", "Species")
  # fix taxa names
  tax_table(my_phyloseq)[,colnames(tax_table(my_phyloseq))] <- gsub(tax_table(my_phyloseq)[,colnames(tax_table(my_phyloseq))],pattern="[a-z]__",replacement="")
  return(my_phyloseq)
}

# from_phyloseq_web 
veganotu = function(physeq) {
  require(vegan)
    OTU = otu_table(physeq)
    if (taxa_are_rows(OTU)) {
        OTU = t(OTU)
    }
    return(as(OTU, "matrix"))
}
```



Analysis of fungal sequence data derived from the same DNA extractions as those used in our previous bacterial analysis ([Elliott et al 2014](http://dx.doi.org/10.1007/s10531-014-0684-8)). The analysis for that paper can be found [here](http://davidelliott.github.io/kalahari_bacteria/kalahari_BSC_bacteria.html). 

This analysis in R follows the upstream preparation in file kf_analysis.sh and QIIME-notes.txt. 

# Revision notes
This analysis was updated after reviewer feedback as follows:


* Use latest version of UNITE for taxonomy (upstream from this script, documented in QIIME-notes.txt)
* Due to better taxonomy in the new UNITE, the custom approach to taxonomy is removed
* Slightly alter the DESeq analysis to work on the exact same subsetted objects as was used for the ADONIS - this makes a bit more logical sense and does not affect the results much. Specifically, the depth analysis is performed separately on the open grass sites (=cyanobacterial biocrust areas), and canopy places.
* Use volcano plots to display the DESeq2 results. This includes using a more sophisticated approach to OTU selection for comparisons
* Run the fungi data through the pipeline used for the bacteria paper - in separate script "kalahari_fungi_using_bacteria_pipeline.Rmd"

# Import data and setup

## Fungi
```{r, import_fungi, echo=FALSE,message=FALSE, warning=FALSE}
library("phyloseq"); library("ggplot2"); library("scales"); library("grid"); library("reshape2"); library(RColorBrewer); library("DESeq2"); library("xtable"); library("vegan"); library("ape"); library(qiime2R); library(EnhancedVolcano);library(ggpubr);library(gridExtra)

# BiocManager::install('EnhancedVolcano')

biom <- import_biom("data/fungi/otutab.biom",taxaPrefix="X")
map <- import_qiime_sample_data("data/fungi/map.txt")
map$zone2line <- sub(pattern="_",replacement="\n",x=map$zone)
expt <- qza_to_phyloseq(features = "data/fungi/feature-table-1.qza",
                taxonomy = "data/fungi/unite-v9-dynamic-vsearch-classifications.qza")
expt <- merge_phyloseq(expt, map)
expt <- tidy_phyloseq(expt)

# add bacterial phylum abundance to the environmental data
bac_phylum <- read.csv("data/bacteria/otu_table_bacteria_phylum_rel.csv",na.strings = "")
bac_phylumt = setNames(data.frame(t(bac_phylum[,-1])), bac_phylum[,1])
colnames(bac_phylumt) <- gsub(pattern = "-",replacement = ".",x = colnames(bac_phylumt))
sample_data(expt) <- cbind(sample_data(expt), bac_phylumt[row.names(map),])

# Re-formatting and adding new names to sample data for plots
sample_data(expt)$depth2 <- "0-1 cm"
sample_data(expt)$depth2[sample_data(expt)$depth == 1] <- "1-2 cm"
samples <- as.character(sample_data(expt)$X.SampleID)
sample_data(expt)$samples <- as.factor(substr(samples, 1, nchar(samples)-1))
sample_data(expt)$zc_depth <- paste(sample_data(expt)$zone_code,sample_data(expt)$depth,sep="")
# Identify the dry season and wet season sampling
sample_data(expt)$season <- "D"
sample_data(expt)$season[sample_data(expt)$month=="Mar"] <- "W"

print(expt)

# Prepare some phyloseq and other objects for later use
expt.rel <- transform_sample_counts(expt, function(x) 100*x/(sum(x)))
# check that the above worked properly: sample_sums(expt.rel) > yes - all samples add up to 100

expt.rel.class <- tax_glom(expt.rel,taxrank="Class")
expt.rel.phylum <- tax_glom(expt.rel,taxrank="Phylum")

expt.rel.phylum.all <- merge_samples(expt.rel.phylum,"ReversePrimer")
expt.rel.phylum.all <- transform_sample_counts(expt.rel.phylum.all, function(x) x/48)
expt.rel.phylum.type <- merge_samples(expt.rel.phylum,"type")
expt.rel.phylum.type <- transform_sample_counts(expt.rel.phylum.type, function(x) x/24)
expt.rel.phylum.veg <- merge_samples(expt.rel.phylum,"zone")
expt.rel.phylum.veg <- transform_sample_counts(expt.rel.phylum.veg, function(x) x/12)

taxa_names(expt.rel.phylum.all) <- tax_table(expt.rel.phylum.all)[,2]
sample_names(expt.rel.phylum.all) <- "overall"
taxa_names(expt.rel.phylum.type) <- tax_table(expt.rel.phylum.type)[,2]
taxa_names(expt.rel.phylum.veg) <- tax_table(expt.rel.phylum.veg)[,2]

map.df <- as.data.frame(sample_data(expt))
map.df$n.sequences <- colSums(otu_table(expt))[row.names(map.df)]

```


## Bacteria
Bacterial data are from Elliott et al (2014)

```{r bac_import, echo=FALSE,message=FALSE}

# set input filenames
otufilename <- "data/bacteria/otu_table.biom"
mapfilename <- "data/bacteria/Elliott_2188B_map_sff.txt"

biom <- import_biom(otufilename,taxaPrefix="X")
map <- import_qiime_sample_data(mapfilename=mapfilename)
bacteria <- merge_phyloseq(biom, map)
# Fix ranks. 
colnames(tax_table(bacteria)) = c("Kingdom", "Phylum", "Class", "Order", "Family", "Genus", "Species")
# fix taxa names
tax_table(bacteria)[,colnames(tax_table(bacteria))] <- gsub(tax_table(bacteria)[,colnames(tax_table(bacteria))],pattern="[a-z]__",replacement="")

# Add new names to sample data for plots
sample_data(bacteria)$depth2 <- "0-1 cm"
sample_data(bacteria)$depth2[sample_data(bacteria)$depth == 1] <- "1-2 cm"

# Remove Mitochondria and Chloroplast. No mitochondria found.
bacteria <- subset_taxa(bacteria,Class!="Chloroplast")

print(bacteria)
```


```{r set_formatting,echo=FALSE}
# set default ggplot theme
theme_set(theme_bw(base_size = 12, base_family = ""))
alpha = 0.75
# define colour scales
color_var = unique(c(get_variable(expt, "type"),get_variable(expt, "zone_code")))

color_pal = brewer.pal(length(color_var), "Set1")
names(color_pal) <- color_var
color_pal["T"] <- "#F781BF" # don't like yellow
color_pal["0-1 cm"] <- color_pal["crust"]
color_pal["1-2 cm"] <- color_pal["subsoil"]

fill_pal <- color_pal
fill_pal["subsoil"] <- "white"
fill_pal["open"] <- "white"
fill_pal["1-2 cm"] <- fill_pal["subsoil"]
fill_pal["0-1 cm"] <- fill_pal["crust"]

depth_fill <- color_pal[c(7,8)]

# define shape scale
shape_var = c(levels((get_variable(expt, "type"))), "0-1 cm", "1-2 cm")
shape_scale = c(19,24,19,24)
names(shape_scale) = c(shape_var)

# copy parts of the colour scales for convenience:
zone_colours <- color_pal[c("AG",      "PG",      "S",       "T")]

# zone/depth shapes and colours - added during proofing stage
zcd <- sort(unique(map.df$zc_depth))

zcd_shapes <- c(21,21,22,22,23,23,24,24)
zcd_colours <- color_pal[c("AG","AG","PG","PG","S","S","T","T")]
zcd_fill <- zcd_colours
zcd_fill[c(2,4,6,8)] <- "#FFFFFF"
zcd_solid <- c(TRUE,FALSE,TRUE,FALSE,TRUE,FALSE,TRUE,FALSE)

names(zcd_shapes) = zcd
names(zcd_colours) = zcd
names(zcd_fill) = zcd
names(zcd_solid) = zcd

zone_shapes <- zcd_shapes[c(1,3,5,7)]
names(zone_shapes) <- names(zone_colours) 
```

# Main Tables & Figures

## Figure 1. Mixed tree, shrub and grass rangeland at the study site. 
Extensive bare soil patches as shown here between grasses are common and support the development of biocrusts. Photographs of individual vegetation types can be found in Elliott et al (2014). (not shown here)

## Figure 2. Fungal class abundance by vegetation zone and depth in Kalahari arenosols, southern Botswana. 

```{r class_plots_alt2,echo=FALSE,message=FALSE, fig.width=8, fig.height=4, warning=FALSE}

TopNOTUs <- names(sort(taxa_sums(expt.rel.class), TRUE)[1:10])
expt.rel.class.topN <- prune_taxa(TopNOTUs, expt.rel.class)
c <- tax_table(expt.rel.class.topN)[,"Class"]
p <- tax_table(expt.rel.class.topN)[,"Phylum"]
taxa_names(expt.rel.class.topN) <- paste(p,"__",c,sep="")
my_phyloseq <- expt.rel.class.topN

# get the OTU tables
OTUs.table <- data.frame(otu_table(my_phyloseq))
OTUs.t <- t(OTUs.table)

sample_data(my_phyloseq)$zone_type <- paste(sample_data(my_phyloseq)$zone_code,sample_data(my_phyloseq)$type,sep="_")

sample_data(my_phyloseq)$sample <- row.names(sample_data(my_phyloseq))
sample_data <- sample_data(my_phyloseq)[,c("sample","zone","zone_code","zone2line","cover","type","depth2","month")]

# combine OTU and sample data 
OTUs_and_env_data <- data.frame(sample_data,OTUs.t)

OTUs_and_env_data.melt <- melt(OTUs_and_env_data)
# re-name the OTU's that had names changed by melting
OTUs_and_env_data.melt$variable <- gsub(pattern="X",replacement="",x=OTUs_and_env_data.melt$variable)
# Add linebreak in the OTU name (the OTUs were strategically named a few lines up)
OTUs_and_env_data.melt$variable <- gsub(pattern="__",replacement="\n",x=OTUs_and_env_data.melt$variable)
OTUs_and_env_data.melt$variable <- gsub(pattern="_cls_Incertae_sedis",replacement="",x=OTUs_and_env_data.melt$variable)

# plot the OTUs

zone_plot2 <- (ggplot(data=OTUs_and_env_data.melt,aes(x=zone_code,y=value, fill=depth2))
      + facet_wrap(~variable,nrow = 2) 
      + theme_bw(base_size = 9, base_family = "") 
      + theme(aspect.ratio=1)
      + scale_x_discrete(name="nearby vegetation") 
      + scale_y_continuous(name="% abundance",trans = 'log10',breaks=c(0.1,1,10,100))  
      + geom_boxplot()  
      + theme(axis.text.x = element_text(angle = -90, vjust = 0.4))
      + scale_fill_manual(values=fill_pal,name="Depth") )
```

Boxes represent the interquartile range (IQR), and error bars extend to the most extreme values within 1.5 * IQR of the box. Median values are shown as a line within the box and outliers are shown as black spots. Sample coding: AG=annual grass, PG=perennial grass, S=shrub, T=tree. The ten most abundant classes are shown, accounting for `r sum(taxa_sums(expt.rel.class.topN)/nsamples(expt.rel.class.topN)) ` of sequence reads.
```{r class_plots_alt_show,echo=FALSE,message=FALSE, fig.width=8, fig.height=4, warning=FALSE}
# display the figure. Output to PDF is done later.
zone_plot2
```

## Figure 3. Correspondence analysis of the fungal community in Kalahari arenosols, southern Botswana constrained by vegetation and depth. 
Markers indicate the fungal communities of individual samples. Filled markers indicate soil surface/biocrust (0-1 cm depth), open markers indicate subsoil communities (1-2 cm depth). Sample coding: AG=annual grass, PG=perennial grass, S=shrub, T=tree.

```{r ord, echo=FALSE}
expt.ord <- expt
ord <- ordinate(expt.ord,method = "CCA",distance = "bray",formula = ~zone+type)
p.ord <- plot_ordination(expt.ord, ord, type = "sites", color = "zone_code", shape="zone_code") + geom_point(alpha=0.7,aes(fill=zc_depth)) + theme(aspect.ratio=1) + geom_text(aes(label = season),show.legend = FALSE,vjust="bottom", size=2, nudge_y = -0.1)

# Format the points to identify depth, vegetation, and sampling month
p.ord <- p.ord + scale_colour_manual(values=zone_colours, name="Vegetation") + 
scale_fill_manual(values=zcd_fill, name="Vegetation", guide="none") + 
scale_shape_manual(values=zone_shapes, name="Vegetation") 

# display the figure. Output to PDF is done later.
p.ord

```


## checking assumptions for ADONIS2 permanova test:

```{r adnois2_setup,message=FALSE,echo=TRUE,warning=FALSE,cache=TRUE}
  # ADONIS to be done separately for open and canopy areas, at OTU/species rank.
  
  # grass areas
  expt.open <- subset_samples(expt, cover=="open")
  dist.open = phyloseq::distance(expt.open, "bray")
  
  # canopy areas
  expt.canopy <- subset_samples(expt, cover=="canopy")
  dist.canopy = phyloseq::distance(expt.canopy, "bray")
 
  # Checking for homogeneity of group dispersions to ensure that dispersion within groups does not skew the PERMANOVA test.

  ## Calculate multivariate dispersions
  mod.open.zone <- betadisper(dist.open,group = sample_data(expt.open)$zone) # need to check the ordering of data matches up - done - it does match
  mod.open.month <- betadisper(dist.open,group = sample_data(expt.open)$month)
  mod.open.type <- betadisper(dist.open,group = sample_data(expt.open)$type)
  mod.canopy.zone <- betadisper(dist.canopy,group = sample_data(expt.canopy)$zone)
  mod.canopy.month <- betadisper(dist.canopy,group = sample_data(expt.canopy)$month)
  mod.canopy.type <- betadisper(dist.canopy,group = sample_data(expt.canopy)$type)
  
  ## Perform tests
  anova(mod.open.zone)
  anova(mod.open.month)
  anova(mod.open.type)
  anova(mod.canopy.zone)
  anova(mod.canopy.month)
  anova(mod.canopy.type)
  
  # The plots and further analyses below were used to check further

  ## Plot the groups and distances to centroids on the
  ## first two PCoA axes
plot(mod.open.zone)
plot(mod.open.month)
plot(mod.open.type)
plot(mod.canopy.zone)
plot(mod.canopy.month)
plot(mod.canopy.type)
  
  ## Draw a boxplot of the distances to centroid for each group
boxplot(mod.open.zone)
boxplot(mod.open.month)
boxplot(mod.open.type)
boxplot(mod.canopy.zone)
boxplot(mod.canopy.month)
boxplot(mod.canopy.type)
  
  ## Permutation test for F
permutest(mod.open.zone, pairwise = TRUE, permutations = 99)
permutest(mod.open.month, pairwise = TRUE, permutations = 99)
permutest(mod.open.type, pairwise = TRUE, permutations = 99)
permutest(mod.canopy.zone, pairwise = TRUE, permutations = 99)
permutest(mod.canopy.month, pairwise = TRUE, permutations = 99)
permutest(mod.canopy.type, pairwise = TRUE, permutations = 99)

```


## Table 1. Permutational analysis of variance to identify significant differences in fungal community structure with respect to vegetation zone, depth, and season in Kalahari arenosols of southern Botswana. 
This table shows results for the community data at OTU (97 % similarity). The “all samples” analysis was also carried out at multiple taxonomic ranks with similar results as shown in supplementary data. 

(table needs to be made manually from the results below)

```{r adonis2,message=FALSE,echo=FALSE,warning=FALSE,cache=TRUE,}

  NMDS = ordinate(expt.open, "NMDS", dist.open)
  print(adonis2(dist.open ~ zone + depth + month , as(sample_data(expt.open), "data.frame")) )

  NMDS = ordinate(expt.canopy, "NMDS", dist.canopy)
  print(adonis2(dist.canopy ~ zone + depth + month , as(sample_data(expt.canopy), "data.frame")) )

```

(see also Note S1)

Use differential abundance analysis to identify which fungi are associated with the cyanobacterial biocrust containing areas (open grass interspaces), or canopy areas.
```{r DESeq_new1, echo=FALSE, message=FALSE, warning=FALSE}
# Identify fungi which are associated with the biocrust areas (open grass interspaces)

# set a variable for the desired log change cut-off
log_threshold <- 2 # set at 4-fold

# use all the full experiment data in this first test
expt.deseq <- expt

# add 1 to the whole OTU table, to avoid infinity problems
# this will reduce the fold changes slightly
otu_table(expt.deseq) <- otu_table(expt.deseq)+1

coverdds = phyloseq_to_deseq2(expt.deseq, ~ cover)
coverdds = DESeq(coverdds, test="Wald",fitType="parametric")

# plotDispEsts(coverdds,main="model check")
res = results(coverdds, cooksCutoff = FALSE, alpha = 0.05)
alpha = 0.05
deseq_cover_results <- res

# From the initial DESeq analysis on all samples with respect to cover type
# identify the OTUs which are associated with the open areas (grass interspaces) which harbour cyanobacterial biocrusts
# these will be the ones where Log change is >threshold and P<0.05

index1 <- deseq_cover_results$log2FoldChange>log_threshold # grass interspaces
index2 <- deseq_cover_results$log2FoldChange<(log_threshold*-1) # tree/shrub canopy
index3 <- deseq_cover_results$padj<0.05

cat("grass interspace associated OTUs (place where cyanobacterial biocrusts are)")
(biocrust_associated_OTUs <- row.names(deseq_cover_results)[index1&index3])
cat("canopy associated OTUs (place lacking cyanobacterial biocrusts)")
(canopy_associated_OTUs <- row.names(deseq_cover_results)[index2&index3])

```

Check to see if any of the biocrust associated fungi preferentially occupy the surface or sub-surface soil. Also perform the same check on the fungi which are associated with the non-biocrust areas (canopy).
```{r DESeq_new2, echo=FALSE, message=FALSE, warning=FALSE}

# Now perform the differential abundance analysis on the biocrust associated OTUs to see if any of them preferentially occupy the surface or sub-surface soil

expt.biocrust <- prune_taxa(biocrust_associated_OTUs , expt)

  expt.deseq <- expt.biocrust
  # add 1 to the whole OTU table, to avoid infinity problems
  # this will reduce the fold changes slightly
  otu_table(expt.deseq) <- otu_table(expt.deseq)+1
  
  bc_depth_dds = phyloseq_to_deseq2(expt.deseq, ~ depth)
  bc_depth_dds = DESeq(bc_depth_dds, test="Wald",fitType="parametric")

  # plotDispEsts(bc_depth_dds,main="model check")

  res = results(bc_depth_dds, cooksCutoff = FALSE, alpha = 0.05)
  alpha = 0.05

  deseq_biocrust_by_depth_results <- res

# same again for canopy assocaited taxa
  
expt.canopy <- prune_taxa(canopy_associated_OTUs , expt)

  expt.deseq <- expt.canopy

  # add 1 to the whole OTU table, to avoid infinity problems
  # this will reduce the fold changes slightly
  otu_table(expt.deseq) <- otu_table(expt.deseq)+1
  
  depthdds = phyloseq_to_deseq2(expt.deseq, ~ depth)
  depthdds = DESeq(depthdds, test="Wald",fitType="parametric")

  # plotDispEsts(coverdds,main="model check")

  res.depth = results(depthdds, cooksCutoff = FALSE, alpha = 0.05)
  alpha = 0.05
  # sigtab_depth = res.depth
  
  deseq_canopy_by_depth_results <- res.depth

# for the depth results we are interested in both the OTUs that have significantly differential abundance AND those that do not
# To highlight this in figure, we can label those OTUS - which means most of them, but excluding the ambiguous few in the middle
# Criteria:
  # P<0.05 & fold change >log_threshold
  # P>0.05 & fold change <log_threshold
  subsurface_OTUs.open <- deseq_biocrust_by_depth_results$padj<0.05 & abs(deseq_biocrust_by_depth_results$log2FoldChange)>log_threshold
  cosmopolitan_OTUs.open <- deseq_biocrust_by_depth_results$padj>0.05 & abs(deseq_biocrust_by_depth_results$log2FoldChange)<log_threshold

  wanted_labels_index.open <- subsurface_OTUs.open | cosmopolitan_OTUs.open
  
  subsurface_OTUs.canopy <- deseq_canopy_by_depth_results$padj<0.05 & abs(deseq_canopy_by_depth_results$log2FoldChange)>log_threshold
  cosmopolitan_OTUs.canopy <- deseq_canopy_by_depth_results$padj>0.05 & abs(deseq_canopy_by_depth_results$log2FoldChange)<log_threshold

  wanted_labels_index.canopy <- subsurface_OTUs.canopy | cosmopolitan_OTUs.canopy
```

Create volcano plots to graphically show the results of the differential abundance analyses

```{r volcano_plots_new, echo=FALSE, message=FALSE, warning=FALSE}
# as suggested by reviewer
# using guidance from https://bioconductor.org/packages/devel/bioc/vignettes/EnhancedVolcano/inst/doc/EnhancedVolcano.html
# plotting the results of deseq_cover_results + deseq_biocrust_by_depth_results

annotation_open <- data.frame(
   x = c(-4,0,4),
   y = c(15,15,15),
   label = c("More abundant\n0-1 cm depth","Similarly abundant\nthroughout depths", "More abundant\n1-2 cm depth")
)

annotation_cover <- data.frame(
   x = c(-4.4,4.4),
   y = c(41,41),
   label = c("More abundant in\ncanopy areas\n(these compared by\ndepth in fig 5B)", "More abundant in\ngrass inter-spaces\n(these compared by\ndepth in fig 5A)")
)

annotation_canopy<- data.frame(
   x = c(-4,0,4),
   y = c(12,12,12),
   label = c("More abundant\n0-1 cm depth", "Similarly abundant\nthroughout depths","More abundant\n1-2 cm depth")
)

## Cover results - this plot determines which OTUs are associated to biocrust
OTU_IDs <- rownames(deseq_cover_results)
TAX_abbreviation <- as.character(substr(x = tax_table(expt)[rownames(deseq_cover_results),"Class"] , start = 0, stop = 3))
OTU_TAX <- paste(OTU_IDs,TAX_abbreviation)
OTU_TAX_labels.cover <- substr(x = OTU_TAX , start = 5, stop = 12)

volcano_cover <- EnhancedVolcano(deseq_cover_results,
    lab = OTU_TAX_labels.cover,
    x = 'log2FoldChange',
    y = 'pvalue',
    pCutoff = 0.05,
    FCcutoff = log_threshold,
    #title = 'B. Differential abundance of fungal OTUs by cover type',
    #subtitle = '(areas of grass inter-spaces vs areas under shrubs and trees)',
    pointSize = 3.5,
    labSize = 3.5,
    legendLabels=c('Not significant','Log (base 2) FC','p<0.05',
      'p<0.05 & Log(2) change>2'),
    drawConnectors = TRUE,
    arrowheads = FALSE,
    max.overlaps = 20,
    widthConnectors = 0) + geom_text(data=annotation_cover, aes( x=x, y=y, label=label), size=4 , fontface="italic") 


## depth results. This plot shows which of the biocrust associated taxa are differentially abundant by depth
# create custom names for OTU labelling

OTU_IDs <- rownames(deseq_biocrust_by_depth_results)
TAX_abbreviation <- as.character(substr(x = tax_table(expt)[rownames(deseq_biocrust_by_depth_results),"Class"] , start = 0, stop = 3))
OTU_TAX <- paste(OTU_IDs,TAX_abbreviation)
OTU_TAX_labels.depth <- substr(x = OTU_TAX , start = 5, stop = 12)


volcano_depth <- EnhancedVolcano(deseq_biocrust_by_depth_results,
    lab = OTU_TAX_labels.depth,
    x = 'log2FoldChange',
    y = 'pvalue',
    pCutoff = 0.05,
    FCcutoff = log_threshold,
   # title = 'A. Differential abundance of fungal OTUs by depth',
   # subtitle = '(areas of cyanobacterial biocrust of grass inter-spaces)',
    pointSize = 3.5,
    labSize = 3.5,
   selectLab = OTU_TAX_labels.depth[wanted_labels_index.open],
    legendLabels=c('Not significant','Log (base 2) FC','p<0.05',
      'p<0.05 & Log(2) change>2'),
    drawConnectors = TRUE,
    arrowheads = FALSE,
    max.overlaps = 20,
    widthConnectors = 0) + geom_text(data=annotation_open, aes( x=x, y=y, label=label), size=4 , fontface="italic") 


## depth results. This plot shows which of the biocrust associated taxa are differentially abundant by depth
# create custom names for OTU labelling

OTU_IDs <- rownames(deseq_canopy_by_depth_results)
TAX_abbreviation <- as.character(substr(x = tax_table(expt)[rownames(deseq_canopy_by_depth_results),"Class"] , start = 0, stop = 3))
OTU_TAX <- paste(OTU_IDs,TAX_abbreviation)
OTU_TAX_labels.depth <- substr(x = OTU_TAX , start = 5, stop = 12)


volcano_depth_canopy <- EnhancedVolcano(deseq_canopy_by_depth_results,
    lab = OTU_TAX_labels.depth,
    x = 'log2FoldChange',
    y = 'pvalue',
    pCutoff = 0.05,
    FCcutoff = log_threshold,
   # title = 'A. Differential abundance of fungal OTUs by depth',
   # subtitle = '(areas of cyanobacterial biocrust of grass inter-spaces)',
    pointSize = 3.5,
    labSize = 3.5,
   selectLab = OTU_TAX_labels.depth[wanted_labels_index.canopy],
    legendLabels=c('Not significant','Log (base 2) FC','p<0.05',
      'p<0.05 & Log(2) change>2'),
    drawConnectors = TRUE,
    arrowheads = FALSE,
    max.overlaps = 20,
    widthConnectors = 0) + geom_text(data=annotation_canopy, aes( x=x, y=y, label=label), size=4 , fontface="italic") 

# create custom names for OTU labelling

deseq_cover_plot <- volcano_cover + theme(legend.position="none") + labs(title = NULL, subtitle = NULL,caption = NULL) + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())
deseq_biocrust_depth_plot <- volcano_depth + theme(legend.position="none") + labs(title = NULL, subtitle = NULL,caption = NULL) + labs(tag = "A") + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())
deseq_canopy_depth_plot <- volcano_depth_canopy + theme(legend.position="none") + labs(title = NULL, subtitle = NULL,caption = NULL) + labs(tag = "B") + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) 
deseq_volcano_legend <- as_ggplot(get_legend(volcano_depth))

# cover plot
deseq_cover_plot + labs(title="all taxa by cover type") 
# depth plots
(deseq_biocrust_depth_plot <- deseq_biocrust_depth_plot + labs(subtitle="inter-space associated taxa by depth"))
(deseq_canopy_depth_plot <- deseq_canopy_depth_plot + labs(subtitle="canopy associated taxa by depth"))


# Output to PDF also
pdf(file = "figures/Figure 4 [volcano-plot-cover].pdf", width=9, height=6)
grid.newpage()
pushViewport(viewport(layout = grid.layout(nrow = 2, ncol = 1, heights = c(6,1)))) 
print(deseq_cover_plot, vp = viewport(layout.pos.row = 1))
print(deseq_volcano_legend, vp = viewport(layout.pos.row = 2))
dev.off()

pdf(file = "figures/Figure 5 [volcano-plots-depth].pdf", width=7, height=9)
grid.newpage()
pushViewport(viewport(layout = grid.layout(nrow = 3, ncol = 1, heights = c(4,4,1)))) 
print(deseq_biocrust_depth_plot, vp = viewport(layout.pos.row = 1))
print(deseq_canopy_depth_plot, vp = viewport(layout.pos.row = 2))
print(deseq_volcano_legend, vp = viewport(layout.pos.row = 3))
dev.off()
```



```{r deseq_tables,echo=FALSE}
# the deseq analyses produced multiple outputs depending on the comparisons made
# here they are adjusted and combined for eventual output

#### First build a master table with all the OTU information known
# bigtab = as.data.frame(tax_table(expt))
# changed this to add the taxonomy in later

# merge the deseq results tables and suffix with identification of the originating test

# all OTUs were first compared to see if they are more abundant in canopy or open areas
# This is stored in deseq_cover_results and should include all the OTUs
deseq_cover_results <- as.data.frame(deseq_cover_results)
colnames(deseq_cover_results) <- paste(colnames(deseq_cover_results),".cover",sep="")

# Next, OTUs were compared for abundance differences by depth - this was focused on the OTUs which were more abundant 
# in either canopy or open areas - therefore these analyses do not contain all the OTUs
# These results are stored in deseq_canopy_by_depth_results and deseq_biocrust_by_depth_results
# The following 2 commands merge the 3 results tables together into a table called sigtabs
sigtabs <- merge(as.data.frame(deseq_canopy_by_depth_results),as.data.frame(deseq_biocrust_by_depth_results),by.x=0,by.y=0,all.x=T,all.y=T,suffixes=c(".canopy.depth",".open.depth")) 
sigtabs <- merge(as.data.frame(deseq_cover_results),as.data.frame(sigtabs),by.y="Row.names",by.x=0,all.x=T,all.y=T)

# Start to build a big table of results for export as supplementary data
bigtab <- sigtabs

# set row names as OTU names and remove the rownames column
row.names(bigtab) <- bigtab$Row.names
bigtab <- bigtab[,-1]

# Create dataframes of OTU abundance for output to files later
rel_otu <- cbind(tax_table(expt.rel),otu_table(expt.rel))
expt.zcdepth <- merge_samples(expt.rel,group = "zc_depth")
expt.zcdepth <- transform_sample_counts(expt.zcdepth, function(x) x/6)
rel_otu_by_factor <- cbind(tax_table(expt.zcdepth),t(otu_table(expt.zcdepth)))
rel_otu_by_factor <- as.data.frame(rel_otu_by_factor )

# Add a column for the total abundance. We will also sort the output by that
rel_otu_by_factor[,c("AG0","AG1","PG0","PG1","S0","S1","T0","T1")] <- sapply(rel_otu_by_factor[,c("AG0","AG1","PG0","PG1","S0","S1","T0","T1")],as.numeric)
rel_otu_by_factor$Total <- rowSums(rel_otu_by_factor[,c("AG0","AG1","PG0","PG1","S0","S1","T0","T1")], na.rm=TRUE)
# Divide the total by 8 to give the average abundance per sample across the whole study (because the summed abundances were per factor)
rel_otu_by_factor$Total <- rel_otu_by_factor$Total/8
# sum(rel_otu_by_factor$Total) # it does come to 100 %

# Add the abundance by factor into the bigtab dataframe
bigtab <- merge(rel_otu_by_factor,bigtab,by.x=0,by.y=0) # the number 0 specifies the row names.

# sort bigtab by the overall abundance of OTUs across whole study
bigtab <- bigtab[order(bigtab$Total,decreasing = T),]

# set row names as OTU names and change the row.names column into a numeric index that can be used for sorting
row.names(bigtab) <- bigtab$Row.names
colnames(bigtab)[1] <- "OTU_n"
bigtab[,"OTU_n"] <- as.numeric(substr(x = bigtab[,1] , start = 5, stop=9))

# make a list of OTUs that are associated by depth
OTU_depth_associated.open <- row.names(bigtab[(!is.na(bigtab$padj.open.depth)) & (bigtab$padj.open.depth<0.05),])
OTU_depth_associated.canopy <- row.names(bigtab[(!is.na(bigtab$padj.canopy.depth)) & (bigtab$padj.canopy.depth<0.05),])

```


## Mantel comparison of fungal and bacterial communities
Compare community distance matrix of fungi with the bacteria distance matrix using data from Elliott et al (2014). Two tests are run:

- open zones (grass interspace)
- canopy zones (under tree and shrub)

```{r mantel, echo=FALSE}

# get the bacteria OTU table in the same order as the fungi one
otu_table(bacteria) <- otu_table(bacteria)[,sample_names(expt)]

# prepare subset objects for canopy and open areas
fun.open <- subset_samples(expt, cover=="open")
bac.open <- subset_samples(bacteria, cover=="open")
fun.canopy <- subset_samples(expt, cover=="canopy")
bac.canopy <- subset_samples(bacteria, cover=="canopy")

# make distance matrices for open areas
bac.dist.open = phyloseq::distance(bac.open, "bray")
fun.dist.open = phyloseq::distance(fun.open, "bray")

# make distance matrices for canopy areas
bac.dist.canopy = phyloseq::distance(bac.canopy, "bray")
fun.dist.canopy = phyloseq::distance(fun.canopy, "bray")

# Check that the distance matrices are all in the same order
# identical(labels(fun.dist.open),labels(bac.dist.open))
# identical(labels(fun.dist.canopy),labels(bac.dist.canopy))
# yes they are

# run the Mantel tests
mantel(bac.dist.open,fun.dist.open)
mantel(bac.dist.canopy,fun.dist.canopy)

```

# Supplementary Tables & Figures

## Rarefaction
```{r rare,echo=FALSE,fig.width=9, fig.height=6}
exptv <- veganotu(expt)
exptv.individual <- veganotu(expt)
# identical(row.names(exptv),row.names(sample_data(expt)))
row.names(exptv) <- sample_data(expt)$zc_depth 
cols <- as.numeric(as.factor(row.names(exptv)))
rarecurve(exptv,step = 100,label = T,cex=0.6,bty="n",col=cols)
```

## Table S1. Sample data and sample level results
```{r}
rich <- estimate_richness(expt,split=TRUE)
map.df <- cbind(map.df,rich[row.names(map.df),])
```
Various tables are in the [tables](tables/) folder. Table S1 is called sample_data.csv.

## Table S2. Relative abundance of fungal phlya in Kalahari arenosols of southern Botswana. 
Overall results are shown in addition to breakdown by depth and nearby vegetation type. 

```{r manuscript_calcs, results='asis',echo=FALSE, warning=FALSE}
a <- as.data.frame(t(otu_table(expt.rel.phylum.all)))
b <- as.data.frame(t(otu_table(expt.rel.phylum.type)))
c <- as.data.frame(t(otu_table(expt.rel.phylum.veg)))

phylum_summary <- cbind(a,b,c)
# order by overall
phylum_summary <- phylum_summary[order(phylum_summary$overall,decreasing = T),] 

print(type="html",xtable(phylum_summary))

```


## Table S3. OTU summary including taxonomy and differential abundance analysis results.

Various tables are in the [tables](tables/) folder. Table S3 is called differential-abundance.csv.


## Note S1. ADONIS test for community differences with respect to vegetation zone, depth, and sampling month.
This analysis was performed using adonis() function of the vegan package, which is described as follows:
> Analysis of variance using distance matrices - for partitioning distance matrices among sources of variation and fitting linear models (e.g., factors, polynomial regression) to distance matrices; uses a permutation test with pseudo-F ratios.

ADONIS test was done using the whole experiment data at all taxonomic ranks.

```{r adonis,message=FALSE,echo=FALSE,warning=FALSE,cache=TRUE}
my_rank = rank_names(expt)[2]
for (my_rank in c(rank_names(expt)[2:5],"OTU")) {
  cat("\n\n__",my_rank,"__\n")
  if(my_rank!="OTU") {
    expt.taxrank <- tax_glom(physeq=expt,taxrank=my_rank)
  } else {
    expt.taxrank <- expt
  }  
  dist = phyloseq::distance(expt.taxrank, "bray")
  # dist = phyloseq::distance(expt, "unifrac") # Unifrac gives similar results
  NMDS = ordinate(expt.taxrank, "NMDS", dist)
  print(adonis2(dist ~ zone + depth + month , as(sample_data(expt.taxrank), "data.frame")) )
}
```



# Extra supporting information not for publication directly (may inform manuscript or support other result)

## Figure xx. OTU richness estimation (Chao1) and diversity index (Shannon) by site and depth in Kalahari arenosols, southern Botswana. 
The mean and standard error are shown (n=6). Sample coding: AG=annual grass, PG=perennial grass, S=shrub, T=tree.

(not using this now in manuscript, just kept here for reference.)

```{r rich, echo=FALSE, warning=FALSE, fig.width=8, fig.height=4,message=FALSE, results='hide'}
p <- plot_richness(expt, x = "zone_code", color = "depth2", shape = "depth2", measures = c("Chao1","Shannon")) 
p2 <- p + stat_summary(fun.data = "mean_se", position = position_dodge(width = 0.2))

p2$layers <- p2$layers[c(-1,-2)]
p2 <- p2 + theme(aspect.ratio=1) + xlab("vegetation")
(p2 <- p2 + scale_colour_discrete(name="Depth") + scale_shape_discrete(name="Depth"))

p2d <- p2$data
p <- plot_richness(expt, x = "zone_code", color = "depth2", shape = "depth2") 

```

## Table xx. Significance of vegetation type, depth, and season on fungal phylum abundance in Kalahari arenosols of southern Botswana. (open areas only)
wilcox is used for factors with =2 levels (depth, month, and zone).

(not using this now in manuscript, just kept here for reference.)

```{r wilcox_open, echo=FALSE, warning=FALSE, results='asis'}
expt.rel.phylum.open <- subset_samples(expt.rel.phylum, cover=="open")
# remove zero observations
keep <- taxa_sums(expt.rel.phylum.open)>0
expt.rel.phylum.open <- prune_taxa(keep,expt.rel.phylum.open)
# construct a matrix to hoold results
mat <- matrix(nrow=ntaxa(expt.rel.phylum.open),ncol=6, dimnames = list(taxa_names(expt.rel.phylum.open),c("zone.W", "zone.p", "depth.W","depth.p","month.W","month.p")))
for(i in 1:ntaxa(expt.rel.phylum.open)) {
  phylum <- as.numeric(otu_table(expt.rel.phylum.open)[i,])
  phylum_name <- taxa_names(expt.rel.phylum.open)[i]
  depth <- get_variable(physeq = expt.rel.phylum.open, varName = "depth")
  zone <- get_variable(physeq = expt.rel.phylum.open, varName = "zone")
  month <- get_variable(physeq = expt.rel.phylum.open, varName = "month")
  # perform the tests
  d <- wilcox.test(phylum ~ depth) 
  z <- wilcox.test(phylum ~ zone) 
  m <- wilcox.test(phylum ~ month) 
  # store the results
  mat[phylum_name,"depth.W"] <- d$statistic
  mat[phylum_name,"depth.p"] <- d$p.value
  mat[phylum_name,"month.W"] <- m$statistic
  mat[phylum_name,"month.p"] <- m$p.value
  mat[phylum_name,"zone.W"] <- z$statistic
  mat[phylum_name,"zone.p"] <- z$p.value
}

row.names(mat) <- tax_table(expt.rel.phylum)[row.names(mat),2]
print(type="html",xtable(mat,digits=c(0,0,6,0,6,0,6)))
```

## Table xx. Significance of vegetation type, depth, and season on fungal phylum abundance in Kalahari arenosols of southern Botswana.
kruskal is used for factors with >2 levels (zone). wilcox is used for factors with =2 levels (depth and month).

(not using this now in manuscript, just kept here for reference.)

```{r kruskalwilcox, echo=FALSE, warning=FALSE, results='asis'}
# construct a matrix to hold results
mat <- matrix(nrow=ntaxa(expt.rel.phylum),ncol=6, dimnames = list(taxa_names(expt.rel.phylum),c("zone.chisq", "zone.p", "depth.W","depth.p","month.W","month.p")))
for(i in 1:ntaxa(expt.rel.phylum)) {
  phylum <- as.numeric(otu_table(expt.rel.phylum)[i,])
  phylum_name <- taxa_names(expt.rel.phylum)[i]
  depth <- get_variable(physeq = expt.rel.phylum, varName = "depth")
  zone <- get_variable(physeq = expt.rel.phylum, varName = "zone")
  month <- get_variable(physeq = expt.rel.phylum, varName = "month")
  # perform the tests
  d <- wilcox.test(phylum ~ depth) 
  z <- kruskal.test(phylum ~ zone) 
  m <- wilcox.test(phylum ~ month) 
  # store the results
  mat[phylum_name,"depth.W"] <- d$statistic
  mat[phylum_name,"depth.p"] <- d$p.value
  mat[phylum_name,"month.W"] <- m$statistic
  mat[phylum_name,"month.p"] <- m$p.value
  mat[phylum_name,"zone.chisq"] <- z$statistic
  mat[phylum_name,"zone.p"] <- z$p.value
}

row.names(mat) <- tax_table(expt.rel.phylum)[row.names(mat),2]
print(type="html",xtable(mat,digits=c(0,3,6,0,6,0,6)))
```


## Richness / diversity statistical tests

Check effect of depth on richness/diversity:
```{r richness_comparison,echo=FALSE,message=FALSE,warning=FALSE}
# richness category comparison
rich <- estimate_richness(expt, split=TRUE)

# join metadata table to richness results
rich.plus <- cbind(rich[as.character(map.df$X.SampleID),],map.df)

wilcox.test(Shannon ~ type, data=rich.plus)
kruskal.test(Shannon ~ zone_code, data=rich.plus)

wilcox.test(Chao1 ~ type, data=rich.plus, exact=F)
kruskal.test(Chao1 ~ zone_code, data=rich.plus)

# Shannon and Chao1 differ by zone - check further:
library("pgirmess")
cat("Post-hoc test - Shannon diversity")
kruskalmc(Shannon ~ zone,data=rich.plus)
cat("Post-hoc test - Chao1 richness")
kruskalmc(Chao1 ~ zone,data=rich.plus)

# compare diversity by depth in the whole data and separately in each zone
wilcox.test(Shannon ~ type, data=rich.plus)
wilcox.test(Shannon ~ type, data=rich.plus, subset = rich.plus$zone_code=="AG")
wilcox.test(Shannon ~ type, data=rich.plus, subset = rich.plus$zone_code=="PG")
wilcox.test(Shannon ~ type, data=rich.plus, subset = rich.plus$zone_code=="S")
wilcox.test(Shannon ~ type, data=rich.plus, subset = rich.plus$zone_code=="T")

# compare richness by depth in the whole data and separately in each zone
wilcox.test(Chao1 ~ type, data=rich.plus,exact = F)
wilcox.test(Chao1 ~ type, data=rich.plus, subset = rich.plus$zone_code=="AG")
wilcox.test(Chao1 ~ type, data=rich.plus, subset = rich.plus$zone_code=="PG", exact=F)
wilcox.test(Chao1 ~ type, data=rich.plus, subset = rich.plus$zone_code=="S")
wilcox.test(Chao1 ~ type, data=rich.plus, subset = rich.plus$zone_code=="T")


# check for correlation between plant succession and richness/diversity
rich.plus$succession <- 0
rich.plus$succession[rich.plus$zone_code=="AG"] <- 1
rich.plus$succession[rich.plus$zone_code=="PG"] <- 2
rich.plus$succession[rich.plus$zone_code=="S"] <- 3
rich.plus$succession[rich.plus$zone_code=="T"] <- 4

cor.test(x=rich.plus$succession,y=rich.plus$Chao1, method="spearman", exact=F)
cor.test(x=rich.plus$succession,y=rich.plus$Shannon, method="spearman", exact=F)
```



```{r sigplots,echo=FALSE, warning=FALSE, eval=FALSE}
# Save a plot of all the significant OTUs (EVAL=FALSE to save processing time. No need to regenerate the plots every time)

for(otu in OTU_depth_associated.open ) {
    this_otu <- cbind( sample_data(expt.rel)[,c("zone","type")] , t(otu_table(expt.rel)[otu,]) )
    png(paste("figures_extra/sig_depth_open_", otu,".png", sep=""),units="mm", res=600 , width = 174, height = 174)
    print(ggplot(this_otu,aes_string(x="zone",y=otu,colour="type")) + geom_boxplot() + geom_point() + scale_y_log10()) 
    dev.off()
}

for(otu in OTU_depth_associated.canopy ) {
    this_otu <- cbind( sample_data(expt.rel)[,c("zone","type")] , t(otu_table(expt.rel)[otu,]) )
    png(paste("figures_extra/sig_depth_canopy_", otu,".png", sep=""),units="mm", res=600 , width = 174, height = 174)
    print(ggplot(this_otu,aes_string(x="zone",y=otu,colour="type")) + geom_boxplot() + geom_point() + scale_y_log10()) 
    dev.off()
}
```

```{r sigplots2,echo=FALSE, warning=FALSE, eval=FALSE}
# Save a plot of all the significant OTUs (EVAL=FALSE to save processing time. No need to regenerate the plots every time)

for(otu in row.names(sigtab_cover.subset) ) {
    this_otu <- cbind( sample_data(expt.rel)[,c("cover","type")] , t(otu_table(expt.rel)[otu,]) )
    png(paste("figures/sig_cover_", otu,".png", sep=""),units="mm", res=600 , width = 174, height = 174)
    print(ggplot(this_otu,aes_string(x="type",y=otu,colour="cover")) + geom_boxplot() + geom_point() + scale_y_log10()) 
    dev.off()
}
```


## Caclulations for manuscript text

```{r manuscript_calcs2,echo=FALSE,warning=FALSE}
cat("Individual abundance of ten most abundant taxonomic classes" )
taxa_sums(expt.rel.class.topN)/nsamples(expt.rel.class.topN)
cat("Combined abundance of ten most abundant taxonomic classes" )
sum(taxa_sums(expt.rel.class.topN)/nsamples(expt.rel.class.topN))
cat("number of fungal sequence reads")
summary(sample_sums(expt))
cat("total number of fungal sequence reads")
sum(sample_sums(expt))

# To work out the number of OTUs that are significant, we can use the data from the volcano plots:
cover_data <- volcano_cover$data
depth_canopy_data <- volcano_depth_canopy$data
depth_open_data <- volcano_depth$data

# OTUs that were significantly associated with open areas (places harbouring cyanobacterial biocrusts)

# OTUs significantly associated with canopy areas, 

# From the grass-interspace OTUs we found xx associated with the soil surface and xx associated with the sub-surface soil. From canopy associated OTUs we found xx associated with the soil surface and xx associated with the sub-surface soil. 


cat("OTUs significantly more abundant in the inter-space areas (places with biocrust, so they are potentially biocrust associated taxa):\n")
index <- intersect (which(cover_data$Sig=="FC_P") , which(cover_data$log2FoldChange>0) )
(interspace_OTUS <- rownames(cover_data[index,]) )
length(interspace_OTUS)

cat("OTUs significantly more abundant in the canopy areas (places without cyanobacterial biocrust):\n")
index <- intersect (which(cover_data$Sig=="FC_P") , which(cover_data$log2FoldChange<0) )
(canopy_OTUS <- rownames(cover_data[index,]) )
length(canopy_OTUS)

cat("Inter-space associated OTUS that are significantly more abundant in surface (0-1 cm):\n")
index <- intersect (which(depth_open_data$Sig=="FC_P") , which(depth_open_data$log2FoldChange<0) )
(open_surface_OTUS <- rownames(depth_open_data[index,]) )
length(open_surface_OTUS)

cat("Inter-space associated OTUS that are significantly more abundant in subsurface (1-2 cm):\n")
index <- intersect (which(depth_open_data$Sig=="FC_P") , which(depth_open_data$log2FoldChange>0) )
(open_subsurface_OTUS <- rownames(depth_open_data[index,]) )
length(open_subsurface_OTUS)

cat("Inter-space associated OTUS that are similarly abundant in surface and subsurface:\n")
index <- intersect (which(depth_open_data$Sig=="NS") , which(abs(depth_open_data$log2FoldChange)<log_threshold ) )
(open_cosmopolitan_OTUS <- rownames(depth_open_data[index,]) )
length(open_cosmopolitan_OTUS)

cat("Canopy associated OTUS that are significantly more abundant in surface (0-1 cm):\n")
index <- intersect (which(depth_canopy_data$Sig=="FC_P") , which(depth_canopy_data$log2FoldChange<0) )
(canopy_surface_OTUS <- rownames(depth_canopy_data[index,]) )
length(canopy_surface_OTUS)

cat("Canopy associated OTUS that are significantly more abundant in subsurface (1-2 cm):\n")
index <- intersect (which(depth_canopy_data$Sig=="FC_P") , which(depth_canopy_data$log2FoldChange>0) )
(canopy_subsurface_OTUS <- rownames(depth_canopy_data[index,]) )
length(canopy_subsurface_OTUS)

cat("Canopy associated OTUS that are similarly abundant in surface and subsurface:\n")
index <- intersect (which(depth_canopy_data$Sig=="NS") , which(abs(depth_canopy_data$log2FoldChange)<log_threshold ) )
(canopy_cosmopolitan_OTUS <- rownames(depth_canopy_data[index,]) )
length(canopy_cosmopolitan_OTUS)




# identify these OTUS in a column of the bigtab, for inclusion in supplementary data

bigtab$identified_association <- "none"
bigtab[interspace_OTUS,"identified_association"] <- "open areas"
bigtab[canopy_OTUS,"identified_association"] <- "canopy areas"
bigtab[open_surface_OTUS,"identified_association"] <- "open areas - surface"
bigtab[open_subsurface_OTUS,"identified_association"] <- "open areas - subsurface"
bigtab[open_cosmopolitan_OTUS,"identified_association"] <- "open areas - cosmopolitan"
bigtab[canopy_surface_OTUS,"identified_association"] <- "canopy areas - surface"
bigtab[canopy_subsurface_OTUS,"identified_association"] <- "canopy areas - subsurface"
bigtab[canopy_cosmopolitan_OTUS,"identified_association"] <- "canopy areas - cosmopolitan"


```


```{r fig.width=12,fig.height=10,echo=FALSE, message=FALSE, warning=FALSE}
# output some tables...

# Sample metadata
write.csv(map.df,file="tables/Data Sheet 1 [sample_data].csv")

# OTU frequency per sample
write.csv(otu_table(expt),"tables/otu_count.csv")

# Deseq differential abundance results, taxonomy
write.csv(bigtab,file="tables/Data Sheet 2 [taxonomy and differential abundance].csv")

# Relative abundance by sample and by factor
write.csv(rel_otu,file="tables/otus_tax_and_rel_abundance.csv")
write.csv(rel_otu_by_factor,file="tables/otus_tax_and_rel_abundance_by_factor.csv")

```

```{r pdfout,echo=FALSE,message=FALSE,results='hide'}
# output pdf figures

fig2 <- zone_plot2
# fig4 <- p2 + theme_bw(base_size = 13, base_family = "") + theme(aspect.ratio=1) + theme(legend.position="bottom")
fig3 <- p.ord + theme_bw(base_size = 13, base_family = "") + theme(aspect.ratio=1) 
#fig3 <- p.ord + theme_bw(base_family = "") + theme(aspect.ratio=1) 

# individually
pdf(file = "figures/Figure 2 [fungal class abundance].pdf", width=8, height=5)
fig2
dev.off()

#pdf(file = "figures/Figure 4.pdf", width=8, height=6)
#fig4
#dev.off()

pdf(file = "figures/Figure 3 [fungal community correspondence analysis].pdf", width=8, height=8)
fig3
dev.off()

```

